---
title: "OBIS Marine Biodata Workshop"
author: "Stace Beaulieu, Ian Brunjes"
date: "2023-04-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Use Case: Interoperable IFCB data product for the CA HAB Bulletin
# aka IFCB Dashboard to OBIS use case with automated classification
## Event core with Occurrence extension
Ian Brunjes (SCCOOS), Stace Beaulieu (WHOI)
Prepared for OBIS IOOS Marine Biological Data Mobilization Workshop April 2023
**This is a prototype for testing purposes only.**
**A protocol is being developed to determine if and when appropriate to submit products from automated classification to OBIS.**
Sponsored by NOAA PCMHAB20 project “Harmful Algal Bloom Community Technology Accelerator”

Main steps in the workflow can be related to the [IFCB workflow diagram](https://raw.githubusercontent.com/hsosik/ifcb-analysis/master/Development/IFCB%20workflow%20chart.png) in the [ifcb-analysis wiki on GitHub](https://github.com/hsosik/ifcb-analysis/wiki)

Step: Classification
- Interpretation for the autoclass scores / transform automated classification into presence/absence
Step: Summarization
- Calculate concentration as number of ROIs classified to a taxon divided by volume analyzed
Next step (not shown / would extend the diagram): Transforming to Darwin Core
- Map resulting data table into Darwin Core table(s)

### Target data product to standardize to Darwin Core:
Concentration of 2 genera of HAB taxa from an IFCB sample(s) (e.g., [here is a sample with autoclass available in HABDAC at Del Mar Mooring](https://ifcb.caloos.org/timeline?dataset=del-mar-mooring&bin=D20210620T221255_IFCB158)

## R Markdown

```{r , echo=FALSE}


```

